---
title: "Data Exercise 1 (PS813)"
author: "Lotem Bassan"
date: "1/27/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

<!-- Typically, we start with a chunk that loads packages (and installs any that are needed) and sets any defaults (e.g. fig.height). 
As written, this script will only run on a machine that already has several packages installed, which is fine for homework, but not it is not portable.
--> 

#Uploading dataset in STATA
I begin by uploading the dataset using the net install command in STATA:
<!-- I could not run this because the data file was not added to this folder.--> 

```{r}
#net install PS813_EX1, from(https://faculty.polisci.wisc.edu/weimer)

##checking PS813_EX1 consistency and verifying not already installed...
##installing into u:\ado\plus\...
##installation complete.

#PS813_EX1 3451

#saveold "\\Desktop\Grad School\courses\PS813\Rwork\PS1.dta"
##(saving in Stata 12 format, which can be read by Stata 11 or 12)
##file \\Desktop\Grad School\courses\PS813\Rwork\PS1.dta saved
```
<!-- Making an R chunk and then using comments is a fine hack, but a chunk with no engine ```{} is simpler and prittier.-->

#Uploading the file in R
I set the working directory in R and upload the .dta file:
```{r}
setwd("~/Desktop/Grad School/courses/PS813/Rwork")
library(foreign)
PS1Data <- read.dta("PS1.dta")
```
<!-- I'm glad you have a project folder for 813. 
If you are using it as an R Studio project, switching (File->Open Project) to PS813 will automatically change your working directory. 
Then your file paths can just be "Rwork/PS1.dta"--More importantly, no one else has your same file paths. 
That is why we want to kill the habbit of using setwd(). When I see setwd() in replication files, I know I am in for hours of pain. Everything should be written from the perspective of your project folder, so that anyone with that folder can run all of the files in it.
The here() function makes paths work. For example, let's say this script is saved in /Rwork but your data are in /data as they should be. here("data/PS1.dta") will always work.-->


#Rough Plot of the data
```{r}
plot(PS1Data$terms, PS1Data$Leg_Act, 
     xlab = "Number of terms in office", 
     ylab= "Legislative activity of state assembly members", 
     cex.lab = 0.8, cex.axis= 0.8, pch=20)
```
<!-- This works, but ggplot is so much more versitile. Little stuff like this can be an chance to practice.
ggplot(PS1Data) + 
aes(x = terms, y = Leg_Act) + 
geom_point() +
labs(x = "Number of terms in office", 
     y = "Legislative activity of state assembly members")
 -->

#Correlation
Compute correlation between variables:
```{r}
cor(PS1Data$terms, PS1Data$Leg_Act)
```

#Regression
Regress Leg_Act on terms:
```{r}
fit <- lm(Leg_Act ~ terms, data = PS1Data)
summary(fit)
```



#Residuals
First, I create a separate variable for the predicted values using the previous model and the predict() function:
```{r}
PS1Data$p_Leg_Act <- predict(fit)
```
<!-- Nicely done keeping it tidy in the data frame. :) -->

Second, I calculate the residuals as the difference beteen the observed andpredicted values of the dependent variable:
```{r}
PS1Data$Residuals <- PS1Data$Leg_Act - PS1Data$p_Leg_Act
```

Subsequently, I plot the residuals against the independent variable:
```{r}
plot(PS1Data$terms, PS1Data$Residuals, 
     xlab = "Number of terms in office", 
     ylab= "Residuals", 
     cex.lab = 0.8, cex.axis= 0.8, pch=20)
```
<!-- The name of a chunk is the name of the figures it produces. 
You could name this {r residuals, fig.height = 3} -->

In order to get the correlation of the observed and predicted values of the dependent variable, I use the cor() function:
```{r}
correlation <- cor(PS1Data$Leg_Act, PS1Data$p_Leg_Act)
correlation
```
The correlation is `r correlation`.

I now square the correlation coefficient compare to the reported $R^2$:
```{r}
correlation*correlation
```
<!-- No need for a chunk! -->
The correlation squared is `r correlation*correlation`.
