---
title: "Data Exercise 1 (PS813)"
author: "Lotem Bassan"
date: "1/27/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

#Uploading dataset in STATA
I begin by uploading the dataset using the net install command in STATA:

```{r}
#net install PS813_EX1, from(https://faculty.polisci.wisc.edu/weimer)

##checking PS813_EX1 consistency and verifying not already installed...
##installing into u:\ado\plus\...
##installation complete.

#PS813_EX1 3451

#saveold "\\Desktop\Grad School\courses\PS813\Rwork\PS1.dta"
##(saving in Stata 12 format, which can be read by Stata 11 or 12)
##file \\Desktop\Grad School\courses\PS813\Rwork\PS1.dta saved
```

#Uploading the file in R
I set the working directory in R and upload the .dta file:
```{r}
setwd("~/Desktop/Grad School/courses/PS813/Rwork")
library(foreign)
PS1Data <- read.dta("PS1.dta")
```

#Rough Plot of the data
```{r}
plot(PS1Data$terms, PS1Data$Leg_Act, 
     xlab = "Number of terms in office", 
     ylab= "Legislative activity of state assembly members", 
     cex.lab = 0.8, cex.axis= 0.8, pch=20)
```

#Correlation
Compute correlation between variables:
```{r}
cor(PS1Data$terms, PS1Data$Leg_Act)
```

#Regression
Regress Leg_Act on terms:
```{r}
fit <- lm(Leg_Act ~ terms, data = PS1Data)
summary(fit)
```



#Residuals
First, I create a separate variable for the predicted values using the previous model and the predict() function:
```{r}
PS1Data$p_Leg_Act <- predict(fit)
```

Second, I calculate the residuals as the difference beteen the observed andpredicted values of the dependent variable:
```{r}
PS1Data$Residuals <- PS1Data$Leg_Act - PS1Data$p_Leg_Act
```

Subsequently, I plot the residuals against the independent variable:
```{r}
plot(PS1Data$terms, PS1Data$Residuals, 
     xlab = "Number of terms in office", 
     ylab= "Residuals", 
     cex.lab = 0.8, cex.axis= 0.8, pch=20)
```

In order to get the correlation of the observed and predicted values of the dependent variable, I use the cor() function:
```{r}
correlation <- cor(PS1Data$Leg_Act, PS1Data$p_Leg_Act)
correlation
```

I now square the correlation coefficient compare to the reported $R^2$:
```{r}
correlation*correlation
```
